---
title: "Assignment 4"
author: "Benjamin Moran \n c3076448 (University of Newcastle)"
date: 'Date Generated: `r Sys.Date()`'
# bibliography: AssX.bib
output:
  pdf_document: default
  html_document:
    df_print: tibble
    fig_caption: yes
subtitle: STAT904 Statistical Consulting
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

#### Plagiarism Statement 
No part of this Assignment has been copied from anyone else, and I have not lent any part of it to anyone else.

# Question 1

Provide a list of 4 books on statistical consulting and details on how to obtain them.

# Question 2

The following data give the readings of the variables High Density Lipoprotein (HDL) total Cholesterol (CHOL) and total Triglyceride (TRI) concentration and the presence or absence of Sinking Pre-Beta taken from 42 subjects.

```{r}
if("Data" %in% list.files()){
chol.dat <- read.delim("Data/cholesterol.txt", header = TRUE, sep = ",")
} else {
chol.dat <- read.delim("Assignments/Ass-4/Data/cholesterol.txt", header = TRUE, sep = ",") 
}

```

The client wishes to investigate the possible relationship between HDL and the three variables CHOL, TRI and SPB.

## (a) List 6 questions that you would ask the client at the first meeting.
## (b) Analyse the data and write a report for the client.

# References
